---
title: "Week 4 Part 1 - CIs & t-tests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 1. Confidence intervals

```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

t.test(otter_length) #default null saying mean otter length=0
```


## 2. T-test (1-sample t-test) - specify null

See a claim that mean otter length is 50in
```{r}
otter_test <- t.test(x=otter_length, mu=50)
otter_test
```

## 3. 2-sample t-test

```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 20)
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x=desert_bighorns, y=sierra_bighorns)
#note: often don't use 95% CI for two-sample t-test
```

## END PART 1

## Part 2: Two-sample t-tests

Create subset of data that only contains info for suv and compact classes of cars
```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))
```

### Exploratory data visualization

Exploratory histograms
```{r}
ggplot(data=compact_suv, aes(x=cty))+
  geom_histogram(bins=15)+
  facet_wrap(~class)
#suv looks pretty normal, compact may be positively skewed
```

### Quantile-quantile plot

closer relationship gets to linear, closer sample observations are to being normally distributed
```{r}
ggplot(data=compact_suv, aes(sample=cty))+
  geom_qq()+
  facet_wrap(~class)
```

### Descriptive statistics

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty),
            sd_city = sd(cty),
            n = n())
#even if don't look normal from histograms, central limit theorem and our sample sizes tell us mean distribution will still be normal
```

### 2-sample t-test

Create vectors containing observations for each sample (city gas mileage)
```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty) #turns dataframe into vector, which is what t-test needs

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

t.test(x=compact_sample, suv_sample)
```
